Unshuffling Permutations

### 2015-07-04 article.tex
reduction

\usepackage[utf8x]{inputenc}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
%\usepackage{llncsdoc}
\usepackage[english]{babel}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usepackage{algorithmic}
\usepackage{float}
\usepackage{todonotes}
\usepackage{subfigure}
\usepackage{enumitem}

@@

\DeclareMathOperator{\LEFT}{\ell}
\DeclareMathOperator{\RIGHT}{r}
\DeclareMathOperator{\QQ}{\mathbb{Q}}
\DeclareMathOperator{\STD}{\mathrm{s}}
\DeclareMathOperator{\SHUFFLE}{\bullet}

@@

\author{%
  Samuele Giraudo\inst{1} \and
  St\'ephane Vialette\inst{1}
}% end author
\institute{%
  Universit\'e Paris-Est, LIGM (UMR 8049), CNRS, UPEM, ESIEE Paris, ENPC,
  F-77454, Marne-la-Vallée, France\\
  \email{samuele.giraudo@univ-mlv.fr} \\
  \email{vialette@univ-mlv.fr}
}% end institute
\date{\today}

@@

We denote by $S_n$ the set of permutations of size $n$ and by $S$ the
set of all permutations. If $\pi$ is a permutation, we denote by $|\pi|$
its size. If $P$ is a subset of $[|\pi|]$, $\pi_{|P}$ is the subword of
$\pi$ consisting in the letters of $\pi$ at the positions specified by
the elements of $P$. Besides, we denote by $\widetilde{\pi}$ (resp.
$\bar \pi$, $\pi^{-1}$) the mirror image (resp. complementary, inverse)
of $\pi$. If $u$ is a word of integers without multiple occurrences of a
same letter, $\STD(u)$ denotes the unique permutation of same size as
$u$ such that for all $i, j \in [|u|]$, $u_i < u_j$ if and only if
$\STD(u)_i < \STD(u)_j$.
\medskip

%%
%% ---- Algebraic issues ----
%%
\section{Algebraic issues}
\label{section:Algebraic issues}
Let use denote by $\QQ[S]$ the linear span of all permutations. We shall
define on this space a kind of shuffle product. For this purpose, we
first define a linear coproduct $\Delta$ on $\QQ[S]$ satisfying for
any permutation $\pi$,
\begin{equation} \label{equ:unshuffling_coproduct}
    \Delta(\pi) =
    \sum_{P_1 \uplus P_2 = [|\pi|]}
    \STD\left(\pi_{|P_1}\right) \otimes \STD\left(\pi_{|P_2}\right).
\end{equation}
We call $\Delta$ the {\em unshuffling coproduct of permutations}. For
instance,
\begin{equation}
    \Delta(213) =
    \epsilon \otimes 213 + 2. 1 \otimes 12 +
    1 \otimes 21 + 2. 12 \otimes 1 + 21 \otimes 1 + 213 \otimes \epsilon,
\end{equation}
\begin{equation}
    \Delta(1234) =
    \epsilon \otimes 1234 + 4. 1 \otimes 123 + 6. 12 \otimes 12 +
    4. 123 \otimes 1 + 1234 \otimes \epsilon,
\end{equation}
\begin{equation}\begin{split} \label{equ:example_unshuffling_coproduct}
    \Delta(1432) & =
    \epsilon \otimes 1432 + 3. 1 \otimes 132 + 1 \otimes 321 +
    3. 12 \otimes 21 \\ & + 3. 21 \otimes 12 + 3. 132 \otimes 1 +
    321 \otimes 1 + 1432 \otimes \epsilon.
\end{split}\end{equation}
\medskip

As any coproduct, $\Delta$ leads to the definition of a product obtained
by duality in the following way. From~\eqref{equ:unshuffling_coproduct},
for any permutation $\pi$, we have
\begin{equation}
    \Delta(\pi) =
    \sum_{\sigma, \nu \in S} \lambda_{\sigma, \nu}^\pi \;
    \sigma \otimes \nu,
\end{equation}
where the $\lambda_{\sigma, \nu}^\pi$ are nonnegative integers. Now,
by definition of duality, the dual product of $\Delta$ denoted by
$\SHUFFLE$ is a linear binary product on $\QQ[S]$. It satisfies, for any
permutations
$\sigma$ and $\nu$,
\begin{equation}
    \sigma \SHUFFLE \nu =
    \sum_{\pi \in S}
    \lambda_{\sigma, \nu}^\pi \; \pi.
\end{equation}
We call $\SHUFFLE$ the {\em shuffle product of permutations}. For
instance,
\begin{equation}\begin{split} \label{equ:example_shuffle_product}
    12 \SHUFFLE 21 & =
    1243 + 1324 + 2.1342 + 2.1423 + 3.1432 + 2134 + 2.2314 \\
    & + 3.2341 + 2413 + 2.2431 + 2.3124 + 3142 + 3.3214 + 2.3241 \\
    & + 3421 + 3.4123 + 2.4132 + 2.4213 + 4231 + 4312.
\end{split}\end{equation}
Observe that the coefficient $3$ of the permutation $1432$
in~\eqref{equ:example_shuffle_product} comes from the fact that the
coefficient of the tensor $12 \otimes 21$ is $3$
in~\eqref{equ:example_unshuffling_coproduct}.
\medskip

We say that a permutation $\pi$ {\em appears} in the shuffle
$\sigma \SHUFFLE \nu$ of two permutations $\sigma$ and $\nu$ if the
coefficient $\lambda_{\sigma, \nu}^\pi$ is different from zero.
Besides, a permutation $\pi$ is a {\em square} if there is a permutation
$\sigma$ such that $\pi$ appears in $\sigma \SHUFFLE \sigma$. In this
case, we say that $\sigma$ is a {\em square root} of $\pi$. Equivalently,
$\pi$ is a square with $\sigma$ as square root if and only if in the
expansion of $\Delta(\pi)$, there is a tensor $\sigma \otimes \sigma$
with a nonzero coefficient.
\medskip

\begin{proposition} \label{prop:shuffle_associative_commutative}
    The shuffle product $\SHUFFLE$ of permutations is associative and
    commutative.
\end{proposition}
\begin{proof}
    To prove the associativity of $\SHUFFLE$, it is convenient to show
    that its dual coproduct $\Delta$ is coassociative, that is
    \begin{equation}
        (\Delta \otimes I) \Delta = (I \otimes \Delta) \Delta,
    \end{equation}
    where $I$ denote the identity map. This strategy relies on the fact
    that a product is associative if and only if its dual coproduct is
    coassociative. For any permutation $\pi$, we have
    \begin{equation} \begin{split}
    \label{equ:shuffle_associative_commutative}
        (\Delta \otimes I) \Delta(\pi) & =
        (\Delta \otimes I)
        \sum_{P_1 \uplus P_2 = [|\pi|]}
        \STD\left(\pi_{|P_1}\right) \otimes \STD\left(\pi_{|P_2}\right) \\
        & =
        \sum_{P_1 \uplus P_2 = [|\pi|]}
        \Delta\left(\STD\left(\pi_{|P_1}\right)\right)
        \otimes I\left(\STD\left(\pi_{|P_2}\right)\right) \\
        & =
        \sum_{P_1 \uplus P_2 = [|\pi|]}
        \sum_{Q_1 \uplus Q_2 = [|P_1|]}
        \STD\left(\STD\left(\pi_{|P_1}\right)_{|Q_1}\right)
        \otimes
        \STD\left(\STD\left(\pi_{|P_1}\right)_{|Q_2}\right)
        \otimes \STD\left(\pi_{|P_2}\right) \\
        & =
        \sum_{P_1 \uplus P_2 \uplus P_3 = [|\pi|]}
        \STD\left(\pi_{|P_1}\right) \otimes
        \STD\left(\pi_{|P_2}\right) \otimes
        \STD\left(\pi_{|P_3}\right).
    \end{split} \end{equation}
    An analogous computation shows that $(I \otimes \Delta) \Delta(\pi)$
    is equal to the last member of~\eqref{equ:shuffle_associative_commutative},
    whence the associativity of $\SHUFFLE$.
    \smallskip

    Finally, to prove the commutativity of $\SHUFFLE$, we shall show
    that $\Delta$ is cocommutative, that is for any permutation $\pi$,
    if in the expansion of $\Delta(\pi)$ there is a tensor
    $\sigma \otimes \nu$ with a coefficient $\lambda$, there is in the
    same expansion the tensor $\nu \otimes \sigma$ with the same
    coefficient $\lambda$. Clearly, a product is associative if and only
    if its dual coproduct is coassociative. Now, from the
    definition~\eqref{equ:unshuffling_coproduct} of $\Delta$, one
    observes that if the pair $(P_1, P_2)$ of subsets of $[|\pi|]$
    contributes to the coefficient of
    $\STD\left(\pi_{|P_1}\right) \otimes \STD\left(\pi_{|P_2}\right)$,
    the pair $(P_2, P_1)$ contributes to the coefficient of
    $\STD\left(\pi_{|P_2}\right) \otimes \STD\left(\pi_{|P_1}\right)$.
    This shows that $\Delta$ is cocommutative and hence, that $\SHUFFLE$
    is commutative.
\end{proof}
\medskip

Proposition~\ref{prop:shuffle_associative_commutative} implies that
$\QQ[S]$ endowed with the unshuffling coproduct $\Delta$ is a
coassociative cocommutative coalgebra, or in an equivalent way, that
$\QQ[S]$ endowed with the shuffle product $\SHUFFLE$ is an associative
commutative algebra.
\medskip

\begin{proposition} \label{prop:endomorphisms}
    The three linear maps
    \begin{equation}
        \phi_1, \phi_2, \phi_3 : \QQ[S] \to \QQ[S]
    \end{equation}
    linearly sending a permutation $\pi$ to, respectively,
    $\widetilde{\pi}$, $\bar \pi$, and $\pi^{-1}$ are endomorphims of
    associative algebras.
\end{proposition}
\begin{proof}
    To prove that, for $j = 1, 2, 3$, $\phi_j$ is a morphism of
    associative algebras, we have to prove that for all permutations
    $\sigma$ and $\nu$,
    \begin{equation}
        \phi_j(\sigma \SHUFFLE \nu) =
        \phi_j(\sigma) \SHUFFLE \phi_j(\nu).
    \end{equation}
    We shall proceed here as we have done during the proof of
    Proposition~\ref{prop:shuffle_associative_commutative} by showing
    that $\phi_j$ is a morphism of coalgebras, that is
    \begin{equation}
        \Delta \phi_j = (\phi_j \otimes \phi_j) \Delta.
    \end{equation}
    In the sequel, $\pi$ is a permutation.
    \smallskip

    If $P$ is a set of indexes of letters of $\pi$, we denote by
    $\widetilde{\pi}$ the set $\{|\pi| - i  + 1 : i \in P\}$. Now since
    this operation $\widetilde{\,}$ define a bijection on the set of the
    subsets of $[|\pi|]$ and the standardisation operation commutes with
    the mirror operation on words, we have
    \begin{equation} \begin{split}
        \Delta(\phi_1(\pi))
        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
        \STD\left(\phi_1(\pi)_{|P_1}\right)
        \otimes \STD\left(\phi_1(\pi)_{|P_2}\right) \\
        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
        \STD\left(\widetilde{\pi}_{|P_1}\right)
        \otimes \STD\left(\widetilde{\pi}_{|P_2}\right) \\
        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
        \STD\left(\widetilde{\pi}_{|\widetilde{P_1}}\right)
        \otimes \STD\left(\widetilde{\pi}_{|\widetilde{P_2}}\right) \\
        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
        \widetilde{\STD\left(\pi_{|P_1}\right)}
        \otimes \widetilde{\STD\left(\pi_{|P_2}\right)} \\
        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
        \phi_1\left(\STD\left(\pi_{|P_1}\right)\right)
        \otimes \phi_1\left(\STD\left(\pi_{|P_2}\right)\right) \\
        & = (\phi_1 \otimes \phi_1) \Delta(\pi).
    \end{split} \end{equation}
    This shows that $\phi_1$ is a morphism of coalgebras and hence, that
    $\phi_1$ is a morphism of associative algebras.
    \smallskip

    Next, since by definition of the complementation operation on
    permutations, for any permutation $\tau$ and any indexes $i$ and $k$,
    we have $\tau_i < \tau_j$ if and only if $\bar \tau_i > \bar \tau_j$,
    we have
    \begin{equation} \begin{split}
        \Delta(\phi_2(\pi))
        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
        \STD\left(\phi_2(\pi)_{|P_1}\right)
        \otimes \STD\left(\phi_2(\pi)_{|P_2}\right) \\
        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
        \STD\left(\bar \pi_{|P_1}\right)
        \otimes \STD\left(\bar \pi_{|P_2}\right) \\
        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
        \phi_2\left(\STD\left(\pi_{|P_1}\right)\right)
        \otimes \phi_2\left(\STD\left(\pi_{|P_2}\right)\right) \\
        & = (\phi_2 \otimes \phi_2) \Delta(\pi).
    \end{split} \end{equation}
    This shows that $\phi_2$ is a morphism of coalgebras and hence, that
    $\phi_2$ is a morphism of associative algebras.
    \smallskip

    Finally, for any permutation $\tau$, if $P$ is a set of indexes of
    letters of $\tau$, we denote by $P_\tau^{-1}$ the set
    $\{\tau_i : i \in P\}$. Since the map sending a subset $P$ of
    $[|\pi|]$ to $P_\pi^{-1}$ is a bijection and
    $\STD\left(\pi_{|P}\right)^{-1} = \STD\left(\pi^{-1}_{|P_\pi^{-1}}\right)$,
    we have
    \begin{equation} \begin{split}
        \Delta(\phi_3(\pi))
        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
        \STD\left(\phi_3(\pi)_{|P_1}\right)
        \otimes \STD\left(\phi_3(\pi)_{|P_2}\right) \\
        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
        \STD\left(\pi^{-1}_{|P_1}\right)
        \otimes \STD\left(\pi^{-1}_{|P_2}\right) \\
        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
        \STD\left(\pi^{-1}_{|{P_1}_\pi^{-1}}\right)
        \otimes \STD\left(\pi^{-1}_{|{P_2}_\pi^{-1}}\right) \\
        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
        \STD\left(\pi_{|P_1}\right)^{-1}
        \otimes \STD\left(\pi_{|P_2}\right)^{-1} \\
        & = \sum_{P_1 \uplus P_2 = [|\pi|]}
        \phi_3\left(\STD\left(\pi_{|P_1}\right)\right)
        \otimes \phi_3\left(\STD\left(\pi_{|P_2}\right)\right) \\
        & = (\phi_3 \otimes \phi_3) \Delta(\pi).
    \end{split} \end{equation}
    This shows that $\phi_3$ is a morphism of coalgebras and hence, that
    $\phi_3$ is a morphism of associative algebras.
\end{proof}
\medskip

\begin{proposition} \label{prop:square_stability}
    Let $\pi$ be a square permutation and $\rho$ be a square root of
    $\pi$. Then,
    \begin{enumerate}[label={\it (\roman*)},fullwidth]
        \item \label{item:square_stability_1}
        the permutation $\widetilde{\pi}$ is a square and
        $\widetilde{\rho}$ is one of its square roots;
        \item \label{item:square_stability_2}
        the permutation $\bar \pi$ is a square and $\bar \rho$ is one of
        its square roots;
        \item \label{item:square_stability_3}
        the permutation $\pi^{-1}$ is a square and $\rho^{-1}$ is one of
        its square roots.
    \end{enumerate}
\end{proposition}
\begin{proof}
    All statements~\ref{item:square_stability_1},
    \ref{item:square_stability_2}, and~\ref{item:square_stability_3} are
    direct consequences of Proposition~\ref{prop:endomorphisms}. Indeed,
    since $\pi$ is a square permutation and $\rho$ is a square root of
    $\pi$, by definition $\pi$ appears in the product
    $\rho \SHUFFLE \rho$. Now, by Proposition~\ref{prop:endomorphisms},
    for all $j = 1, 2, 3$, since $\phi_j$ is a morphism of associative
    algebras from $\QQ[S]$ to $\QQ[S]$, $\phi_j$ commutes with the
    shuffle product of permutations $\SHUFFLE$. Hence, in particular,
    one has
    \begin{equation}
        \phi_j(\rho \SHUFFLE \rho) =
        \phi_j(\rho) \SHUFFLE \phi_j(\rho).
    \end{equation}
    Then, since $\pi$ appears in $\rho \SHUFFLE \rho$, $\phi_j(\pi)$
    appears in $\phi_j(\rho) \SHUFFLE \phi_j(\rho)$. This
    implies~\ref{item:square_stability_1}, \ref{item:square_stability_2},
    and~\ref{item:square_stability_3}.
\end{proof}
\medskip

@@

\begin{proposition}
Let $\pi \in S_{2n}$.
Then,
$\pi$ is the union of two order-isomorphic pattern
if and only if
there exists a perfect containment-free mathing
$\mathcal{M}$ of $G_\pi$ such that
$\LEFT(\mathcal{M}) \simeq \RIGHT(\mathcal{M})$.
\end{proposition}

\begin{proposition}
Deciding whether a permutation is the union of two
order-isomorphic pattern is \NPC.
\end{proposition}

@@

\begin{claim}
  \label{claim:no nu_1 - nu_1 edge}
  There is no $(\nu_1, \nu_2)$-edge in $\mathcal{M}$.
\end{claim}

\begin{proof}[of Claim~\ref{claim:no nu_1 - nu_1 edge}]
Suppose, aiming at a contradiction that there exists
a $(\nu_1, \nu_2)$-edge in $\mathcal{M}$.

\begin{tikzpicture}
  [scale=0.25]
  % nu_1
  \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,9.8) grid (5.2,15.2);
  \foreach \x/\y in {0/10,1/11,2/12,3/13,4/14,5/15} {
      \draw [fill=black] (\x,\y) circle (0.1);
  }
  % nu_4
  \draw[step=1cm,black!30,ultra thin,fill=black!10] (4.8,-0.2) grid (10.2,5.2);
  \foreach \x/\y in {5/5,6/4,7/3,8/2,9/1,10/0} {
      \draw [fill=black] (\x,\y) circle (0.1);
  }
  % nu'_1
  \draw[step=1cm,black!30,ultra thin,fill=black!10] (9.8,14.8) grid (15.2,20.2);
  \foreach \x/\y in {10/15,11/16,12/17,13/18,14/19,15/20} {
      \draw [fill=black] (\x,\y) circle (0.1);
  }
  % edges
  \draw [line width=1pt]
  (1,11) .. controls +(0,2) and +(0,1) .. (2,12);
  \draw [line width=1pt]
  (3,13) .. controls +(0,6) and +(0,16) .. (8,2);
\end{tikzpicture}
\begin{tikzpicture}
  [scale=0.25]
  % nu_1
  \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,9.8) grid (5.2,15.2);
  \foreach \x/\y in {0/10,1/11,2/12,3/13,4/14,5/15} {
      \draw [fill=black] (\x,\y) circle (0.1);
  }
  % nu_4
  \draw[step=1cm,black!30,ultra thin,fill=black!10] (4.8,-0.2) grid (10.2,5.2);
  \foreach \x/\y in {5/5,6/4,7/3,8/2,9/1,10/0} {
      \draw [fill=black] (\x,\y) circle (0.1);
  }
  % nu'_1
  \draw[step=1cm,black!30,ultra thin,fill=black!10] (9.8,14.8) grid (15.2,20.2);
  \foreach \x/\y in {10/15,11/16,12/17,13/18,14/19,15/20} {
      \draw [fill=black] (\x,\y) circle (0.1);
  }
  % edges
  \draw [line width=1pt]
  (1,11) .. controls +(0,2) and +(0,1) .. (2,12);
  \draw [line width=1pt]
  (3,13) .. controls +(0,6) and +(0,16) .. (8,2);
\end{tikzpicture}
\qed
\end{proof}

\begin{claim}
There is at least one $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
\end{claim}

>>>>

\usepackage[utf8x]{inputenc}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{graphicx}
%\usepackage{llncsdoc}
\usepackage[english]{babel}
\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\usepackage{algorithmic}
\usepackage{float}
\usepackage{todonotes}
\usepackage{subfigure}

@@

\DeclareMathOperator{\LEFT}{\ell}
\DeclareMathOperator{\RIGHT}{r}

\newtheorem{fact}{Fact}

@@

\author{%
  Samuele Giraudo \and
  St\'ephane Vialette
}% end author
\institute{%
  Universit\'e Paris-Est, LIGM (UMR 8049), CNRS, UPEM, ESIEE Paris, ENPC,
  F-77454, Marne-la-Vallée, France\\
  \email{samuele.giraudo@univ-mlv.fr} \\
  \email{vialette@univ-mlv.fr}
}% end institute
\date{\today}

@@

@@

\begin{proposition}
  \label{proposition:matching}
  Let $\pi \in S_{2n}$.
  Then,
  $\pi$ is the union of two order-isomorphic patterns
  if and only if
  there exists a perfect containment-free mathing
  $\mathcal{M}$ of $G_\pi$ such that
  $\LEFT(\mathcal{M}) \simeq \RIGHT(\mathcal{M})$.
\end{proposition}

\begin{proposition}
  \label{proposition:hardness}
  Deciding whether a permutation is the union of two
  order-isomorphic patterns is \NPC.
\end{proposition}

\begin{figure}[t!]
  \centering
  \begin{tikzpicture}[
    scale=.95,
    >=stealth',
    shorten >=1pt,
    main node/.style={align=center},
    cell/.style={draw,ultra thick,fill=black!5},
    structure link/.style={line width=1.5pt},
    pattern link/.style={line width=1.5pt,black!20},
    monotone/.style={->,thick}
    ]  
    % edges
    % pi 1 - pi 2
    \draw [pattern link] (8.5,6) .. controls +(0,1) and +(0,2) .. (10.5,5);
    % pi 2 - sigma 2
    \draw [pattern link] (10.5,5) .. controls +(0,1) and +(0,3) .. (11.5,4);
    % 1
    \draw [structure link] (0.5,11) .. controls +(0,2) and +(0,1) .. (2.5,12);
    \node (claim edge 1) at (1.5,13) {Claim\ref{}};
    % 2
    \draw [structure link] (1.5,2) .. controls +(0,1) and +(0,3) .. (6.5,1);
    \node (claim edge 1) at (4,3.5) {Claim\ref{}};
    % 3
    \draw [structure link] (3.5,9) .. controls +(0,3) and +(0,1.5) .. (7.5,10);
    \node (claim edge 1) at (5.5,11.5) {Claim\ref{}};
    % 4
    \draw [structure link] (5.5,8) .. controls +(0,1) and +(0,7) .. (9.5,3);
    \node (claim edge 1) at (8.5,8.5) {Claim\ref{}};
    % sigma 1 - pi 1
    \draw [structure link] (4.5,6) .. controls +(0,-2) and +(0,-1) .. (8.5,5);
    \node (claim edge 1) at (6.5,4) {Claim\ref{}};
    % nodes
    % 1
    \draw [cell] (0,10) -- (1,10) -- (1,11) -- (0,11) -- cycle;
    \draw [monotone] (0,10) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu_1$};
    %\draw [draw,ultra thick,fill=black!5] (0,10) -- (1,10) -- (1,11) -- (0,11) -- cycle;
    \draw [cell] (2,11) -- (3,11) -- (3,12) -- (2,12) -- cycle;
    \draw [monotone] (2,11) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu'_1$};
    % 2
    \draw [cell] (1,1) -- (2,1) -- (2,2) -- (1,2) -- cycle;
    \draw [monotone] (1,2) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu_2$};
    \draw [cell] (6,0) -- (7,0) -- (7,1) -- (6,1) -- cycle;
    \draw [monotone] (6,1) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu'_2$};
    % 3
    \draw [cell] (3,8) -- (4,8) -- (4,9) -- (3,9) -- cycle;
    \draw [monotone] (3,8) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu_3$};
    \draw [cell] (7,9) -- (8,9) -- (8,10) -- (7,10) -- cycle;
    \draw [monotone] (7,9) -- ++(1,1) node [midway,fill=white,fill=black!5] {$\nu'_3$};
    % 4
    \draw [cell] (5,7) -- (6,7) -- (6,8) -- (5,8) -- cycle;
    \draw [monotone] (5,8) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu_4$};
    \draw [cell] (9,2) -- (10,2) -- (10,3) -- (9,3) -- cycle;
    \draw [monotone] (9,3) -- ++(1,-1) node [midway,fill=white,fill=black!5] {$\nu'_4$};
    % sigma 1
    \draw [cell] (4,6) -- (5,6) -- (5,7) -- (4,7) -- cycle;
    \node [main node] (sigma1) at (4.5,6.5) {$\sigma'$};
    % pi 1
    \draw [cell] (8,5) -- (9,5) -- (9,6) -- (8,6) -- cycle;
    \node [main node] (pi1) at (8.5,5.5) {$\pi'$};    
    % pi 2
    \draw [cell] (10,4) -- (11,4) -- (11,5) -- (10,5) -- cycle;
    \node [main node] (sigma2) at (10.5,4.5) {$\pi''$};
    % sigma 2
    \draw [cell] (11,3) -- (12,3) -- (12,4) -- (11,4) -- cycle;
    \node [main node] (pi2) at (11.5,3.5) {$\sigma''$};  
    % grid
    \draw[step=1cm,black!30,ultra thin,fill=black!10] (0,0) grid (12,12);
    % row size
    \foreach \y/\N in {0.5/N_2,1.5/N_2} {
        \node at (-0.5,\y) (R\y) {$\N$};
    }
    % column size
    \foreach \x/\N in 
    {0.5/N_1,1.5/N_2,2.5/N_1,3.5/N_3,4.5/k+2,5.5/N_4,6.5/N_2,7.5/N_3,8.5/n+2,9.5/N_4,10.5/n,11.5/k} {
        \node at (\x,-0.5) (C\x) {$\N$};
    }
  \end{tikzpicture}
  \caption{\label{fig:reduction}}
\end{figure}

@@

\begin{figure}[t!]
  \centering
  \subfigure[A nested $(\nu_1,\nu'_1)$-edge.]{%
    \begin{tikzpicture}
      [
        scale=0.3,
        label/.style={anchor=base}
      ]
      % nu_1
      \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,9.8) grid (5.2,15.2);
      \foreach \x/\y in {0/10,1/11,2/12,3/13,4/14,5/15} {
          \draw [fill=black] (\x,\y) circle (0.1);
      }
      % nu_4
      \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,-0.2) grid (11.2,5.2);
      \foreach \x/\y in {6/5,7/4,8/3,9/2,10/1,11/0} {
          \draw [fill=black] (\x,\y) circle (0.1);
      }
      % nu'_1
      \draw[step=1cm,black!30,ultra thin,fill=black!10] (11.8,15.8) grid (17.2,21.2);
      \foreach \x/\y in {12/16,13/17,14/18,15/19,16/20,17/21} {
          \draw [fill=black] (\x,\y) circle (0.1);
      }
      % labels
      \node [label] (x) at (17,14.5) {$x$};
      % \node [label] (y) at (14,14.5) {$y$};
      \node [label] (a) at (0,8.5) {$a$};
      \node [label] (b) at (8,-1.5) {$b$};
      \node (nu1) at (2.5,7.7) {$\nu_1$};
      \node (nu2) at (8.5,-2.5) {$\nu_2$};
      \node (nu1') at (14.5,13.5) {$\nu'_1$};
      % edges
      \draw [line width=1pt]
      (1,11) .. controls +(0,12) and +(0,4) .. (12,16);
      \draw [line width=1pt]
      (2,12) .. controls +(0,12) and +(0,4) .. (13,17);
      \draw [line width=1pt]
      (3,13) .. controls +(0,12) and +(0,4) .. (14,18);
      \draw [line width=1pt]
      (4,14) .. controls +(0,12) and +(0,4) .. (15,19);
      \draw [line width=1pt]
      (5,15) .. controls +(0,12) and +(0,4) .. (16,20);
      \draw [line width=1pt]
      (0,10) .. controls +(0,8) and +(0,20) .. (8,3);
    \end{tikzpicture}
    \label{subfig:no (nu_1, nu_2)-edge - 2}}
  \qquad
  \subfigure[A nested $(\nu_1,\nu'_1)$-edge.]{%
    \begin{tikzpicture}
      [
        scale=0.3,
        label/.style={anchor=base}
      ]
      % nu_1
      \draw[step=1cm,black!30,ultra thin,fill=black!10] (-0.2,9.8) grid (5.2,15.2);
      \foreach \x/\y in {0/10,1/11,2/12,3/13,4/14,5/15} {
          \draw [fill=black] (\x,\y) circle (0.1);
      }
      % nu_4
      \draw[step=1cm,black!30,ultra thin,fill=black!10] (5.8,-0.2) grid (11.2,5.2);
      \foreach \x/\y in {6/5,7/4,8/3,9/2,10/1,11/0} {
          \draw [fill=black] (\x,\y) circle (0.1);
      }
      % nu'_1
      \draw[step=1cm,black!30,ultra thin,fill=black!10] (11.8,15.8) grid (17.2,21.2);
      \foreach \x/\y in {12/16,13/17,14/18,15/19,16/20,17/21} {
          \draw [fill=black] (\x,\y) circle (0.1);
      }
      % labels
      \node [label] (x) at (3,8.5) {$x$};
      \node [label] (y) at (14,14.5) {$y$};
      \node [label] (a) at (1,8.5) {$a$};
      \node [label] (b) at (8,-1.5) {$b$};
      \node (nu1) at (2.5,7.7) {$\nu_1$};
      \node (nu2) at (8.5,-2.5) {$\nu_2$};
      \node (nu1') at (14.5,13.5) {$\nu'_1$};
      % edges
      \draw [line width=1pt]
      (3,13) .. controls +(0,9) and +(0,4) .. (14,18);
      \draw [line width=1pt]
      (1,11) .. controls +(0,6) and +(0,16) .. (8,3);
    \end{tikzpicture}
    \label{subfig:no (nu_1, nu_2)-edge - 2}}
    \caption{\label{fig:subfig:no (nu_1, nu_2)-edge}%
            Claim~\ref{claim:no (nu_1, nu_2)-edge}.%
            }
  \end{figure}

\begin{claim}
  \label{claim:no (nu_1, nu_2)-edge}
  There is no $(\nu_1, \nu_2)$-edge in $\mathcal{M}$.
\end{claim}

\begin{proof}[of Claim~\ref{claim:no (nu_1, nu_2)-edge}]
  Suppose, aiming at a contradiction that there exists
  a $(\nu_1, \nu_2)$-edge $e = (a, b)$ in $\mathcal{M}$.
  Without loss of generality, assume $(a, b)$ is the leftmost
  $(\nu_1, \nu_2)$-edge in $\mathcal{M}$.
  \begin{itemize}
  \item $a=1$. 
  Since $\mathcal{M}$ is containment-free, it does
  not contain any $(\nu_1, \nu_1)$-edge.
  Frthermore,
  since $N_1 > 2(2N_2 + 2n + 2k + 4) > 
  |\nu_2| + |\nu'_2| + |\nu_3| + |\nu'_3| + |\nu_4| + |\nu'_4|
  + |\sigma'| + |\pi'| + |\pi''| + |\sigma''|$, there exists 
  a $(\nu_1, \nu'_1)$-edge $e' = (x, y)$ in $\mathcal{M}$.
  We have 
  $\LEFT(e) = \mu(a) < \LEFT(e') = \mu(x)$
  and
  $\RIGHT(e') = \mu(b) < \RIGHT(e') = \mu(y)$
  (see Fig.~\ref{fig:reduction}).
  Therefore, since $\mathcal{M}$ is containment-free, it does
  not contain any $(\nu_2, \nu_2)$-edge.
  According to Proposition~\ref{proposition:matching}, this is 
  a contradiciton.
  \end{itemize}
  \qed
\end{proof}

\begin{fact}
  \label{fact:no (nu_2, nu_2)-edge}
  There is no $(\nu_2, \nu_2)$-edge in $\mathcal{M}$.
\end{fact}

\begin{claim}
  \label{fact:one (nu_1, nu'_1)-edge}
  There is at least one $(\nu_1, \nu'_1)$-edge in $\mathcal{M}$.
\end{claim}

\begin{claim}
  \label{fact:no (nu_2, nu'_1)-edge}
  There is no $(\nu_2, \nu'_1)$-edge in $\mathcal{M}$.
\end{claim}

\begin{claim}
  \label{fact:one (nu_2, nu'_2)-edge}
  There is at least one $(\nu_2, \nu'_2)$-edge in $\mathcal{M}$.
\end{claim}

\begin{fact}
  \label{fact:no (nu'_1, nu'_1),(nu_3, nu_3),(sigma' ,sigma'),(nu_4, nu_4)-edge}
  There is 
  no $(\nu'_1, \nu'_1)$-edge, 
  no $(\nu_3, \nu_3)$-edge,
  no $(\sigma', \sigma')$-edge and
  no $(\nu_4, \nu_4)$-edge
  in $\mathcal{M}$.
\end{fact}

\begin{claim}
  \label{fact:no (nu_3, nu'_2)-edge}
  There is no $(\nu_3, \nu'_2)$-edge in $\mathcal{M}$.
\end{claim}

%%
%% ---- Patterns ----
%%
\section{Patterns}
\label{section:Patterns}

\begin{proposition}
  \label{proposition:patterns:lower bound}
  Let $\pi \in S_{n}$.
  Then $\pi$ contains a pattern of size at least 
  $\left\lfloor\sqrt{n-1}/2\right\rfloor$ that
  is the union of two order-isomorphic patterns.
\end{proposition}

\begin{proof}[of Proposition~\ref{proposition:patterns:lower bound}]
  The Erdös-Szekeres theorem \cite{Erdos:Szekeres:1935} 
  states that every permutation of 
  length at least $pq+1$ must contain either the pattern
  $\nearrow_{p+1}$ or the pattern $\searrow_{q+1}$.
  The result now follows from the fact that
  the pattern $\nearrow_{p+1}$ (resp. $\searrow_{q+1}$)
\qed
\end{proof}

\begin{proposition}
  \label{proposition:patterns:upper bound}
  If $2^{\binom{k}{2}} > \binom{n}{2k}\;\binom{2k}{k}$, then there exists
  a permutation of $S_{n}$ 
  that does not contains a pattern
  of length $2k$ that is the union of two
  order-isomorphic patterns.
\end{proposition}
